Powdery mildew yield loss mitigation strategies in mungbean, a meta-analysis.

### 2020-02-23 03_Preliminary_analysis.Rmd
Update preliminary analysis

```{r PA_Libraries, include=FALSE}
library(tidyverse)
library(bomrang)
library(openxlsx)
library(kableExtra)
library(ggplot2)
library(devtools)
library(reshape2)
library(lme4)
library(theme.usq)
source("R/import_data.R")
#source("./getPPD2.R")
```

## Explore and visualise Data

@@

```{r Fungicides}
PM_MB_means %>%
   group_by(fungicide_ai, trial_ref) %>%
   summarise() %>%
   count(sort = TRUE) %>%
   rename(Trials = n) %>%
   ggplot(aes(x = reorder(fungicide_ai, Trials), y = Trials)) +
   xlab("Fungicide active ingredient") +
   ylab("N Trials") +
   geom_col() +
   scale_fill_usq() +
   ggtitle(label = "Number of trials in which the\nspecified fungicide was used") +
   scale_colour_usq() +
   theme_usq() +
   coord_flip()
```

@@

   - Amistar Xtra and Custodia are both contain strobilurin and triazole, however, they because they contain differing dose ratios (inverted) pooling may not be appropriate.

@@

&nbsp;  
&nbsp;

The primary focus of this research is the effect of fungicide on mitigating yield loss to disease. To reduce the affect of location and seasonal influences we will investigate the proportion of grain yield saved, relative to the no spray controls.

First lets investigate the difference in the effect of the type of fungicide on proportion of yield saved. Proportion of yield saved is used in attempt to reduce variation between trials.

@@

```
&nbsp;  
&nbsp;

Lets visualise this in a plot for one and two sprays only.

@@

PM_MB_means %>%
   mutate(fungicide_ai = as.factor(PM_MB_means$fungicide_ai)) %>%
   filter(total_fungicide == 1 | total_fungicide == 2) %>%
   filter(
      fungicide_ai != "pyrazophos" &
         fungicide_ai != "control" &
         fungicide_ai != "benomyl" &
         fungicide_ai != "Acibenzolar-S-methyl") %>%
   mutate(fungicide_ai = factor(fungicide_ai, 
                             levels = c("carbendazim", "pyraclostrobin",
                                        "sulphur", "200 g/L azoxystrobin + 80 g/L cyproconazole",
                                        "tebuconazole", "propiconazole", 
                                        "200 g/L tebuconazole + 120 g/L azoxystrobin"))) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   facet_grid(rows = vars(total_fungicide)) +
   geom_boxplot(aes(fill = fungicide_ai, alpha = 0.5)) +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   labs(y = "Proportion of grain yield saved",
        title = "Grain yield proportion saved\ngrouped by Fungicide type") +
   theme_usq() +
   theme(legend.position = "none") +
   coord_flip()

There is a small yield effect when applying fungicide and it is unclear how much a second spray improves yield saved. Interestingly Sulphur sprays might be worth investigating when sprayed twice.

&nbsp;  
&nbsp;

@@

```{r tebuAndPropi_dose}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole"|
             fungicide_ai == "propiconazole") %>%
   select(trial_ref,
          year,
          location,
          first_sign_disease,
          fungicide_ai,
          dose_ai.ha,
          total_fungicide) %>%
   ggplot(aes(x = as.factor(dose_ai.ha), fill = fungicide_ai)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Total number of treatments for each respective tebuconazole dose") +
   geom_bar() +
   facet_grid(rows = vars(fungicide_ai))+
   scale_fill_usq() +
   scale_colour_usq() +
   theme_usq()
```

@@

There is a large amount of variation in the propiconazole dose, lets inspect the difference in yields for each dose.

```{r PropiDoseEffect}
PM_MB_means %>%
   filter(fungicide_ai == "propiconazole") %>%
   ggplot(aes(x = relevel(as.factor(dose_ai.ha),"62.5"), y = prop_YG)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Yield for each respective propiconazole dose") +
   geom_boxplot() +
   scale_fill_usq() +
   scale_colour_usq() +
   theme_usq()
```
&nbsp;

@@

```{r table_PropiDoses}
table(as.character(PM_MB_means[PM_MB_means$fungicide_ai == "propiconazole",]$trial_ref), 
      PM_MB_means[PM_MB_means$fungicide_ai == "propiconazole",]$dose_ai.ha)
```

@@

```{r PropiDoses_TrialCompar}

boxplot(prop_YG ~ dose_ai.ha, 
        data = PM_MB_means[(PM_MB_means$trial_ref == "mung1112/01" |
                              PM_MB_means$trial_ref == "mung1112/02") &
                              PM_MB_means$fungicide_ai == "propiconazole",])
```

@@

&nbsp;  
&nbsp;

@@

Lets look at the frequency of sprays per fungicide.  
```{r Fungicide_x_nSprays}
table(PM_MB_means$fungicide_ai, PM_MB_means$total_fungicide)

```
&nbsp;  
&nbsp;

@@

```{r spray_relative2FS}
PM_MB_means %>%
   mutate(s1_DfromFS = fungicide_application_1 - first_sign_disease,
          s2_DfromFS = fungicide_application_2 - first_sign_disease,
          s3_DfromFS = fungicide_application_3 - first_sign_disease,
          s4_DfromFS = fungicide_application_4 - first_sign_disease,
          s5_DfromFS = fungicide_application_5 - first_sign_disease) %>%
   gather(key = spray, value = n_days, s1_DfromFS, s2_DfromFS, s3_DfromFS, s4_DfromFS, s5_DfromFS)%>%
   ggplot(aes(x = spray, y = n_days))+
   geom_violin() +
   scale_x_discrete(labels = c("First\nspray", "Second\nspray","Third\nspray","Fourth\nspray","Fifth\nspray"))+
   theme_usq()

&nbsp;  
&nbsp;

@@

```{r Host_genotype}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole"|
             fungicide_ai == "propiconazole")%>%
   group_by(host_genotype, fungicide_ai, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Treatments = n) %>%
   ggplot(aes(x = host_genotype, y = Treatments, fill = fungicide_ai)) +
   xlab("Cultivar") +
   ylab("N Trials") +
   ggtitle(label = "Cultivars used in either tebuconazole or propiconazole trials") +
   geom_col() +
   scale_fill_usq() +
   scale_colour_usq() +
   theme_usq()
```

@@

Does host cultivar significantly impact the yield? Lets look at the volatility in yields to see if any cultivar yields better or has greater tolerance to powdery mildew. Below we plot distributions of mean trial yields for each cultivar.

First lets look at overall yields for each cultivar , where trial is not a factor, for all treatments (spray and no spray).  
```{r yield_vol}
# this could be converted into a shiny app as a chunk. Make each input for the function a shiny option
# Outputs the table and graph
source("R/yield_volatility.r") #function to investigate the volatility in yields
# Volitility is the range between the upper and lower 97.5 and 2.5 % quartiles

@@

YV2 <- yield_volatility(genotype_by_trial = FALSE, control_only = TRUE)
YV2[[2]]
YV2[[3]]

```
Volatility of the cultivars are fairly similar when trial is not considered as a factor (when we inspect plots of all treatments and only spray treatments). Green diamond shows little volatility however this cultivar is only featured in `r length(unique(PM_MB_means[PM_MB_means$host_genotype == "Green Diamond","trial_ref"]))` trials.  
&nbsp;

Lets examine the same plots as above but only for the most commonly used trial site, Hermitage.  
```{r yieldVol_Herm}
YV1_herm <- yield_volatility(genotype_by_trial = FALSE, control_only = FALSE, location = "Hermitage")
YV1_herm[[2]]

@@

Lets take a closer look at Crystal with spray management treatments and without.  
```{r YieldVol_Crystal}
YV1_crystal <- yield_volatility(genotype = "Crystal",control_only = FALSE)
YV1_crystal[[2]]
YV1_crystal <- yield_volatility(genotype = "Crystal",control_only = TRUE)
YV1_crystal[[2]]

@@

```{r YieldVol_JadeControl}
YV3_crystal <- yield_volatility("Jade",control_only = TRUE)
YV3_crystal[[2]]
```

@@

@@

```{r RS_severity}
# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot() +
   #  geom_smooth()+
 #  scale_y_continuous(limits = c(-5, 15)) +
   theme_usq()+
   ggtitle("Powdery mildew variation between different row spacing")

```{r RS_severity_x_trial}
# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot() +
   #  geom_smooth()+
 #  scale_y_continuous(limits = c(-5, 15)) +
   theme_usq()+
   facet_grid(cols = vars(location))+
   ggtitle("Powdery mildew variation between different row spacing")

@@

```{r RS_yield}
# Which row spacing leads to the higher yeild potential
PM_MB_means %>%
   filter(year == 2017|
             year == 2018) %>%
   #   filter(Fungicide == "control") %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = row_spacing,
      colour = location
   )) +
   geom_smooth(method = "gam") +
   geom_jitter(width = 0.01) +
   #   scale_y_continuous(limits = c(-5,15))+
   theme_usq()+
   ggtitle("Grain yield results for different row spacing at three locations")+
   xlab("Row spacing (m)")+
   ylab("Grain yield - tonne per hectare")

@@

lm_rain1 <- read.csv("data/lmInSeasonRainfall_-10.09_30.64.csv",stringsAsFactors = FALSE)
lm_rain2 <- read.csv("data/lmInSeasonRainfall_-10.09_65.100.csv",stringsAsFactors = FALSE)
lm_rain3 <- read.csv("data/lmInSeasonRainfall_10.30_30.64.csv",stringsAsFactors = FALSE)
lm_rain4 <- read.csv("data/lmInSeasonRainfall_10.30_65.100.csv",stringsAsFactors = FALSE)

lm_rain <- rbind(lm_rain1,lm_rain2,lm_rain3,lm_rain4)
```

@@

```{r WhichLModel}
lm_rain[which(lm_rain$lm_pval == min(lm_rain$lm_pval)),]
lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)),]
lm_rain[which(lm_rain$lm_adj_rsquared == max(lm_rain$lm_adj_rsquared)),]

To visualise the model fits we will use a heat map to describe how the models faired in relation to the rainfall windows. log of the p value was used to improve the resolution of small very small p values.  
```{r rainLMpvalue}
ggplot(lm_rain, aes(x = start_day, y = end_day, z = log(lm_pval)))+
   geom_raster(aes(fill = log(lm_pval)), interpolate = TRUE)+
   geom_contour(bins = 15, colour = "white")+
   geom_point(data = lm_rain[which(lm_rain$lm_pval == min(lm_rain$lm_pval)),], aes(x = start_day, y = end_day), colour =" red")+
   scale_fill_distiller(palette ="RdBu", direction = -1)

```{r rainLMrsquared}
ggplot(lm_rain, aes(x = start_day, y = end_day, z = lm_rsquared))+
   geom_raster(aes(fill = lm_rsquared), interpolate = TRUE)+
   geom_contour(bins = 15, colour = "white")+
   geom_point(data = lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)),], aes(x = start_day, y = end_day), colour =" red")+
   scale_fill_distiller(palette ="RdBu") 
```

```{r rainLMadjrsquared}
ggplot(lm_rain, aes(x = start_day, y = end_day, z = lm_adj_rsquared))+
   geom_raster(aes(fill = lm_adj_rsquared), interpolate = TRUE)+
   geom_contour(bins = 15, colour = "white")+
   geom_point(data = lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)),], aes(x = start_day, y = end_day), colour =" red")+
   scale_fill_distiller(palette ="RdBu") 
```

@@

### Season range and first incidence
Mungbean is a summer crop and powdery mildew is requires cool conditions to establish. Therefore the disease can be avoided by planting early in summer or spring, avoiding cooler temperatures at the end of some seasons. Can we visualise when PM is most likely to establish over the calender year.

I'll build a plot of horizontal lines indicating the season length and use a dot on the line to indicating the day of the year first sign was observed for that trial.  
```{r first_incidence}
# Create the data frame for the plot
season_dates <- PM_MB_means %>%
   group_by(trial_ref)%>%
   summarise(Planting_date = unique(planting_date),
             Harvest_date = unique(harvest_date),
             First_sign_PM = unique(first_sign_disease),
             season_length = as.integer(as.Date(unique(harvest_date)) - as.Date(unique(planting_date)))
             )

# function to extract date from the year
# format should be Universal time format
source("R/year2day.r")

season_dates$Planting_day <- unlist(lapply(season_dates$Planting_date,year2day))
season_dates$Disease_day <- unlist(lapply(season_dates$First_sign_PM,year2day))
season_dates$Harvest_day <- season_dates$Planting_day + season_dates$season_length

season_dates[1,"Planting_day"] <- year2day(as.Date("2012-02-28")-7) # planted in december and emerged in February, Assuming due to lack of rain or recording error. Using 7 days prior to emergence day as planting day

first_day_month <- c(0, 31, 59, 90, 120, 151, 181, 211, 243, 273)
axis_labels_date <- c(format(Sys.Date() - year2day(Sys.Date()) + first_day_month[1], "%b-%d"),  # Jan
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[2], "%b-%d"), # Feb
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[3], "%b-%d"), # March
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[4], "%b-%d"), # April
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[5], "%b-%d"), # May
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[6], "%b-%d"), # June
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[7], "%b-%d"), # July
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[8], "%b-%d"), # August
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[9], "%b-%d"), # Sept
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[10], "%b-%d") # October
                         )

@@

# Plot
season_dates %>%
   ggplot()+
   geom_pointrange(aes(x = trial_ref, y = Disease_day, ymin = Planting_day, ymax = Harvest_day))+
   scale_y_continuous(
      limits = c(0, max(season_dates$Planting_day + 
                           season_dates$season_length, 
                        na.rm = TRUE)),
      labels = axis_labels_date,
      breaks = first_day_month
   )+
   geom_point(aes(x = trial_ref, y = Disease_day))+
   coord_flip()+
   ggtitle("Season length and when powdery mildew was first spotted\nin each respective trial")

In most years, regardless of the age of the crop, powdery mildew establishes in March.

&nbsp;  
&nbsp;

#### Effect of Fungicide 
**Fungicide effect on proportional yield gain with respect to time of application**  
First lets look at the quantiles for each fungicide irrespective of other factors.
We will focus on proportional yield gain, compared to no treatment control, as this should partially correct for some spatial and seasonal effects.  
```{r fungicide_proportion_yield}
PM_MB_means %>%
   group_by(fungicide_ai) %>%
   summarise(
      n = length(prop_YG),
      lower_2.5 = quantile(prop_YG, na.rm = T, c(0.025)),
      median = median(prop_YG, na.rm = T),
      mean = mean(prop_YG, na.rm = T),
      upper_97.5 = quantile(prop_YG, na.rm = T, c(0.975))
   ) %>%
   filter(n >= 5) %>% # remove any fungicide groups with less than 5 observations
   arrange(desc(median)) %>%
   kable(caption = "Fungicide effect on proportion of yield gain") %>%
   kable_styling(fixed_thead = T, full_width = T)

# plot table as a box plot
PM_MB_means %>%
   filter(
      fungicide_ai == c(
         "200 g/L tebuconazole + 120 g/L azoxystrobin" ,
         "200 g/L azoxystrobin + 80 g/L cyproconazole",
         "tebuconazole",
         "propiconazole",
         "sulphur"
      )
   ) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   geom_boxplot() +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   #scale_fill_discrete(labels = c("15-7 days before","same day","one day after","8-9 days after","13-19 days after"))+
   labs(x = "Proportion of grain yield saved",
        title = "Grain yield proportion saved grouped by\nfungicide type") +
   theme(
      panel.grid.major = element_line(colour = "grey80"),
      panel.grid.minor = element_line(colour = "grey50"),
      panel.background = element_rect(fill = NA)
   ) +
   guides(fill = guide_legend(title = "Fungicide")) +
   coord_flip()

@@

```{r yield_x_PMseverity}

PM_MB_means %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(x = PM_final_severity, y = grain_yield.t.ha.)) +
   geom_point()+
   geom_smooth(method = "lm")

From the data we have it is difficult to determine if there is an association between powdery mildew severity and grain yield.

>>>>

```{r PA_Libraries, include=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(bomrang)
library(openxlsx)
library(kableExtra)
library(ggplot2)
library(devtools)
library(reshape2)
library(lme4)
library(theme.usq)
source("R/import_data.R")
#source("./getPPD2.R")
theme_set(theme_usq())
```

## Explore and Visualise Data

@@

```{r Fungicides}
PM_MB_means %>%
   group_by(fungicide_ai, trial_ref) %>%
   summarise() %>%
   count(sort = TRUE) %>%
   rename(Trials = n) %>%
   ggplot(aes(x = reorder(fungicide_ai, Trials), y = Trials)) +
   xlab("Fungicide active ingredient") +
   ylab("N Trials") +
   geom_col() +
   scale_fill_usq() +
   ggtitle(label = "Number of trials in which the\nspecified fungicide was used") +
   scale_colour_usq() +
   coord_flip()
```

@@

   - Amistar Xtra and Custodia are both contain strobilurin and triazole, however, because they contain differing dose ratios (inverted) pooling may not be appropriate.

@@

*****

The primary focus of this research is the effect of fungicide on mitigating yield loss to disease. To reduce the effect of location and seasonal influences we will investigate the proportion of grain yield saved relative to the no spray control.

First, let's investigate the difference in the effect of the type of fungicide on proportion of yield saved. Proportion of yield saved is used in attempt to reduce variation between trials.

@@

*****

Let's visualise this in a plot for one and two sprays only.

@@

PM_MB_means %>%
   mutate(fungicide_ai = as.factor(PM_MB_means$fungicide_ai)) %>%
   filter(total_fungicide == 1 | total_fungicide == 2) %>%
   filter(
      fungicide_ai != "pyrazophos" &
         fungicide_ai != "control" &
         fungicide_ai != "benomyl" &
         fungicide_ai != "Acibenzolar-S-methyl"
   ) %>%
   mutate(fungicide_ai = factor(
      fungicide_ai,
      levels = c(
         "carbendazim",
         "pyraclostrobin",
         "sulphur",
         "200 g/L azoxystrobin +\n80 g/L cyproconazole",
         "tebuconazole",
         "propiconazole",
         "200 g/L tebuconazole +\n120 g/L azoxystrobin"
      )
   )) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   facet_grid(rows = vars(total_fungicide)) +
   geom_boxplot(aes(fill = fungicide_ai, alpha = 0.5)) +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   labs(y = "Proportion of grain yield saved",
        title = "Grain yield proportion saved\ngrouped by fungicide type") +
   theme(legend.position = "none") +
   coord_flip()
```

There is a small yield effect when applying fungicide and it is unclear how much a second spray improves yield saved. Interestingly sulphur sprays might be worth investigating when sprayed twice.

*****

@@

```{r tebuAndPropi_dose}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   select(
      trial_ref,
      year,
      location,
      first_sign_disease,
      fungicide_ai,
      dose_ai.ha,
      total_fungicide
   ) %>%
   ggplot(aes(x = as.factor(dose_ai.ha), fill = fungicide_ai)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Total number of treatments for each respective tebuconazole dose") +
   geom_bar() +
   scale_fill_usq() +
   scale_colour_usq() +
   theme_usq()
```

@@

There is a large amount of variation in the propiconazole dose, let's inspect the difference in yields for each dose.

```{r PropiDoseEffect}
PM_MB_means %>%
   filter(fungicide_ai == "propiconazole") %>%
   ggplot(aes(x = relevel(as.factor(dose_ai.ha), "62.5"), y = prop_YG)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Yield for each respective propiconazole dose") +
   geom_boxplot() +
   scale_fill_usq() +
   scale_colour_usq() +
   theme_usq()
```

@@

```{r table_PropiDoses}
table(as.character(PM_MB_means[PM_MB_means$fungicide_ai == "propiconazole", ]$trial_ref),
      PM_MB_means[PM_MB_means$fungicide_ai == "propiconazole", ]$dose_ai.ha)
```

@@

```{r PropiDoses_TrialCompar}
boxplot_usq(prop_YG ~ dose_ai.ha,
            data = PM_MB_means[(PM_MB_means$trial_ref == "mung1112/01" |
                                   PM_MB_means$trial_ref == "mung1112/02") &
                                  PM_MB_means$fungicide_ai == "propiconazole",])
```

@@

*****

@@

Let's look at the frequency of sprays per fungicide.  
```{r Fungicide_x_nSprays}
table(PM_MB_means$fungicide_ai, PM_MB_means$total_fungicide)
```

*****

@@

```{r spray_relative2FS}
PM_MB_means %>%
   mutate(
      s1_DfromFS = fungicide_application_1 - first_sign_disease,
      s2_DfromFS = fungicide_application_2 - first_sign_disease,
      s3_DfromFS = fungicide_application_3 - first_sign_disease,
      s4_DfromFS = fungicide_application_4 - first_sign_disease,
      s5_DfromFS = fungicide_application_5 - first_sign_disease
   ) %>%
   gather(
      key = spray,
      value = n_days,
      s1_DfromFS,
      s2_DfromFS,
      s3_DfromFS,
      s4_DfromFS,
      s5_DfromFS
   ) %>%
   drop_na(n_days) %>% 
   ggplot(aes(x = spray, y = as.numeric(n_days))) +
   geom_violin() +
   scale_x_discrete(
      labels = c(
         "First\nspray",
         "Second\nspray",
         "Third\nspray",
         "Fourth\nspray",
         "Fifth\nspray"
      )
   ) +
   theme_usq()
```

*****

@@

```{r Host_genotype}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(host_genotype, fungicide_ai, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Treatments = n) %>%
   ggplot(aes(x = host_genotype, y = Treatments, fill = fungicide_ai)) +
   xlab("Cultivar") +
   ylab("N Trials") +
   ggtitle(label = "Cultivars used in either tebuconazole or propiconazole trials") +
   geom_col() +
   scale_fill_usq() +
   scale_colour_usq() +
   theme_usq()
```

@@

Does host cultivar significantly impact the yield? Let's look at the volatility in yields to see if any cultivar yields better or has greater tolerance to powdery mildew. Below we plot distributions of mean trial yields for each cultivar.

First lets look at overall yields for each cultivar, where trial is not a factor, for all treatments (spray and no spray).  
```{r yield_vol}
# this could be converted into a shiny app as a chunk. Make each input for the function a shiny option
# Outputs the table and graph
source("R/yield_volatility.r") #function to investigate the volatility in yields
# Volitility is the range between the upper and lower 97.5 and 2.5 % quartiles

@@

YV2 <-
   yield_volatility(genotype_by_trial = FALSE, control_only = TRUE)
YV2[[2]]
YV2[[3]]
```

Volatility of the cultivars are fairly similar when trial is not considered as a factor (when we inspect plots of all treatments and only spray treatments). Green Diamond shows little volatility however this cultivar is only featured in `r length(unique(PM_MB_means[PM_MB_means$host_genotype == "Green Diamond","trial_ref"]))` trials.

*****

Lets examine the same plots as above but only for the most commonly used trial site, Hermitage.  
```{r yieldVol_Herm}
YV1_herm <-
   yield_volatility(
      genotype_by_trial = FALSE,
      control_only = FALSE,
      location = "Hermitage"
   )
YV1_herm[[2]]
```

@@

Lets take a closer look at Crystal with spray management treatments and without.  
```{r YieldVol_Crystal}
YV1_crystal <-
   yield_volatility(genotype = "Crystal", control_only = FALSE)
YV1_crystal[[2]]
YV1_crystal <-
   yield_volatility(genotype = "Crystal", control_only = TRUE)
YV1_crystal[[2]]
```

@@

```{r YieldVol_JadeControl}
YV3_crystal <- yield_volatility("Jade", control_only = TRUE)
YV3_crystal[[2]]
```

@@

*****

@@

```{r RS_severity}
# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot() +
   ggtitle("Powdery mildew variation between different row spacing")
```

```{r RS_severity_x_trial}
# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot() +
   facet_grid(cols = vars(location)) +
   ggtitle("Powdery mildew variation between different row spacing")
```

@@

```{r RS_yield}
# Which row spacing leads to the higher yeild potential
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   #   filter(Fungicide == "control") %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = row_spacing,
      colour = location
   )) +
   geom_smooth(method = "gam") +
   geom_jitter(width = 0.01) +
   ggtitle("Grain yield results for different row spacing at three locations") +
   xlab("Row spacing (m)") +
   ylab("Grain yield - tonnes per hectare")
```

@@

lm_rain1 <-
   read.csv("data/lmInSeasonRainfall_-10.09_30.64.csv",
            stringsAsFactors = FALSE)
lm_rain2 <-
   read.csv("data/lmInSeasonRainfall_-10.09_65.100.csv",
            stringsAsFactors = FALSE)
lm_rain3 <-
   read.csv("data/lmInSeasonRainfall_10.30_30.64.csv",
            stringsAsFactors = FALSE)
lm_rain4 <-
   read.csv("data/lmInSeasonRainfall_10.30_65.100.csv",
            stringsAsFactors = FALSE)

lm_rain <- rbind(lm_rain1, lm_rain2, lm_rain3, lm_rain4)
```

@@

```{r WhichLModel}
lm_rain[which(lm_rain$lm_pval == min(lm_rain$lm_pval)), ]
lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)), ]
lm_rain[which(lm_rain$lm_adj_rsquared == max(lm_rain$lm_adj_rsquared)), ]
```

To visualise the model fits we will use a heat map to describe how the models faired in relation to the rainfall windows. log of the p value was used to improve the resolution of small very small p values.  
```{r rainLMpvalue}
ggplot(lm_rain, aes(x = start_day, y = end_day, z = log(lm_pval))) +
   geom_raster(aes(fill = log(lm_pval)), interpolate = TRUE) +
   geom_contour(bins = 15, colour = "white") +
   geom_point(data = lm_rain[which(lm_rain$lm_pval == min(lm_rain$lm_pval)), ],
              aes(x = start_day, y = end_day),
              colour = " red") +
   scale_fill_distiller(palette = "RdBu", direction = -1) 
```

```{r rainLMrsquared}
ggplot(lm_rain, aes(x = start_day, y = end_day, z = lm_rsquared)) +
   geom_raster(aes(fill = lm_rsquared), interpolate = TRUE) +
   geom_contour(bins = 15, colour = "white") +
   geom_point(data = lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)), ],
              aes(x = start_day, y = end_day),
              colour = " red") +
   scale_fill_distiller(palette = "RdBu") 
```

```{r rainLMadjrsquared}
ggplot(lm_rain, aes(x = start_day, y = end_day, z = lm_adj_rsquared)) +
   geom_raster(aes(fill = lm_adj_rsquared), interpolate = TRUE) +
   geom_contour(bins = 15, colour = "white") +
   geom_point(data = lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)), ],
              aes(x = start_day, y = end_day),
              colour = " red") +
   scale_fill_distiller(palette = "RdBu") 
```

@@

*****

### Season range and first incidence
Mungbean is a summer crop and powdery mildew requires cool conditions to establish.
Therefore the disease can be avoided by planting early in summer or spring, avoiding cooler temperatures at the end of some seasons.
Can we visualise when PM is most likely to establish over the calender year?

I'll build a plot of horizontal lines indicating the season length and use a dot on the line to indicating the day of the year first sign was observed for that trial.  
```{r first_incidence}
# Create the data frame for the plot
season_dates <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(
      Planting_date = unique(planting_date),
      Harvest_date = unique(harvest_date),
      First_sign_PM = unique(first_sign_disease),
      season_length = as.integer(as.Date(unique(harvest_date)) - as.Date(unique(planting_date)))
   )

season_dates$Planting_day <- yday(season_dates$Planting_date)
season_dates$Disease_day <- yday(season_dates$First_sign_PM)
season_dates$Harvest_day <-
   season_dates$Planting_day + season_dates$season_length

season_dates[1, "Planting_day"] <-
   yday(as.Date("2012-02-28") - 7) # planted in december and emerged in February, Assuming due to lack of rain or recording error. Using 7 days prior to emergence day as planting day

first_day_month <- c(0, 31, 59, 90, 120, 151, 181, 211, 243, 273)
axis_labels_date <-
   c(
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[1], "%b-%d"),
      # Jan
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[2], "%b-%d"),
      # Feb
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[3], "%b-%d"),
      # March
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[4], "%b-%d"),
      # April
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[5], "%b-%d"),
      # May
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[6], "%b-%d"),
      # June
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[7], "%b-%d"),
      # July
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[8], "%b-%d"),
      # August
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[9], "%b-%d"),
      # Sept
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[10], "%b-%d") # October
   )

@@

# Plot
season_dates %>%
   ggplot() +
   geom_pointrange(aes(
      x = trial_ref,
      y = Disease_day,
      ymin = Planting_day,
      ymax = Harvest_day
   )) +
   scale_y_continuous(limits = c(
      0,
      max(
         season_dates$Planting_day +
            season_dates$season_length,
         na.rm = TRUE
      )
   ),
   labels = axis_labels_date,
   breaks = first_day_month) +
   geom_point(aes(x = trial_ref, y = Disease_day)) +
   coord_flip() +
   ggtitle("Season length and when powdery mildew was first spotted\nin each respective trial")
```

In most years, regardless of the age of the crop, PM establishes in March.

*****

#### Effect of Fungicide

**Fungicide effect on proportional yield gain with respect to time of application**  
First lets look at the quantiles for each fungicide irrespective of other factors.
We will focus on proportional yield gain, compared to no treatment control, as this should partially correct for some spatial and seasonal effects.  
```{r fungicide_proportion_yield}
PM_MB_means %>%
   group_by(fungicide_ai) %>%
   summarise(
      n = length(prop_YG),
      lower_2.5 = quantile(prop_YG, na.rm = T, c(0.025)),
      median = median(prop_YG, na.rm = T),
      mean = mean(prop_YG, na.rm = T),
      upper_97.5 = quantile(prop_YG, na.rm = T, c(0.975))
   ) %>%
   filter(n >= 5) %>% # remove any fungicide groups with less than 5 observations
   arrange(desc(median)) %>%
   kable(caption = "Fungicide effect on proportion of yield gain") %>%
   kable_styling(fixed_thead = T, full_width = T)

# plot table as a box plot
PM_MB_means %>%
   filter(
      fungicide_ai == c(
         "200 g/L tebuconazole +\n120 g/L azoxystrobin" ,
         "200 g/L azoxystrobin +\n80 g/L cyproconazole",
         "tebuconazole",
         "propiconazole",
         "sulphur"
      )
   ) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   geom_boxplot() +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   #scale_fill_discrete(labels = c("15-7 days before","same day","one day after","8-9 days after","13-19 days after"))+
   labs(x = "Proportion of grain yield saved",
        title = "Grain yield proportion saved grouped by\nfungicide type") +
   theme(
      panel.grid.major = element_line(colour = "grey80"),
      panel.grid.minor = element_line(colour = "grey50"),
      panel.background = element_rect(fill = NA)
   ) +
   guides(fill = guide_legend(title = "Fungicide")) +
   coord_flip()
```

@@

```{r yield_x_PMseverity}
PM_MB_means %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(x = PM_final_severity, y = grain_yield.t.ha.)) +
   geom_point() +
   geom_smooth(method = "lm")
```

From the data that we have, it is difficult to determine if there is an association between powdery mildew severity and grain yield.
